\NeedsTeXFormat{LaTeX2e}[1995/12/01]
\ProvidesFile{ubcsample.tex}[2012/04/07 v1.70 ^^J
 University of British Columbia Thesis]
\documentclass[msc,oneside]{ubcthesis}

% \afterpage{\clearpage} right after a figure.  
\usepackage{afterpage}
\usepackage{float}
\usepackage{longtable} % to define tables that span multiple pages.
\usepackage{pdflscape} % allows for landscape format pages
\usepackage[numbers,sort&compress]{natbib}
\usepackage{psfrag}
\usepackage{graphicx} % for pdf, bitmapped graphics files
\usepackage{amsmath} % assumes amsmath package installed
\usepackage{amssymb}  % assumes amsmath package installed
\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}
%\usepackage{subfigure} % include multiple figures and captions within a single figure environment.
%\usepackage{here}
%\usepackage[subfigure, titles]{tocloft} % This allows you to use math in graph labels
%\usepackage{alltt} % This is useful for including source code from an additional file.
%\usepackage{listings} % has facilities for pretty-printing many languages.

\usepackage[unicode=true,
  linktocpage,
  linkbordercolor={0.5 0.5 1},
  citebordercolor={0.5 1 0.5},
  linkcolor=blue]{hyperref}
%\newcommand\url[1]{\texttt{#1}} % put the URL's in a typewriter font but not link them (comment hyperref first)

\usepackage{setspace}
%\singlespacing
\onehalfspacing
%\doublespacing

\newfloat{Program}{h}{lop}

\institution{The University Of British Columbia}
\faculty{The Faculty of Graduate Studies}
\program{Computer Science}
\institutionaddress{Vancouver}
\previousdegree{B.CS. Hons. Co-op, The University of Manitoba, 2011}
\title{Mapping Environment Properties to Sensory Inputs for Robotic Manipulation}
\author{Daniel Michael Troniak}
\copyrightyear{2013}
\submitdate{\monthname\ \number\year} % The "\ " is required 

\renewcommand\thechapter      {\arabic{chapter}}
\renewcommand\thesection      {\thechapter.\arabic{section}}
\renewcommand\thesubsection   {\thesection.\arabic{subsection}}
\renewcommand\thesubsubsection{\thesubsection.\arabic{subsubsection}}
\renewcommand\theparagraph    {\thesubsubsection.\arabic{paragraph}}
\renewcommand\thesubparagraph {\theparagraph.\arabic{subparagraph}}

\setcounter{tocdepth}{2}
\setcounter{secnumdepth}{2}

\begin{document}

\frontmatter
\maketitle                      %% Mandatory

\begin{abstract}                %% Mandatory -  maximum 350 words
For one to manipulate an object, one must first make judgments about the object such as its mass, shape, size, and function. For example, if I were to pick up an object which was heavier than predicted, such as an overweight suitcase, I would need to detect this fact and adapt my lifting strategy. Thus, to successfully program a robot to perform manipulation tasks, it becomes useful to first equip it with the means to perceive its environment. 

Sensors have been developed that take various measurements such as the amount of force applied to a manipulator or amount of pressure applied to a surface. Gaining abstract understanding of an environment, for example that a suitcase is heavy and its handle is slippery, typically requires combinations of sensor readings to predict accurately. As more sensors are added to the system, however, a cost increase to transmitting and subsequently processing the resulting data is experienced.

In this thesis, I first present existing ways in which touch sensors are used and processed to gain abstract understanding of environments. I then review neuroscience and physiology literature on the current understanding of how the human motor system processes and acts on sensor data. I then present our novel approach to predicting properties of an environment in real-time using a large number of touch sensors, such as force, torque and tactile feedback, mounted to a Barrett WAM robot with attached BarrettHand as the robot performs a prescribed manipulation.

I introduce a novel metric that quantifies the usefulness of any given sensor reading to predicting properties of an environment relevant to the manipulation, including the mass and stiffness of a manipulated block as well as the coefficient of static friction between the block and the table surface. Real-time environment property predictions are made possible by selecting just the top 5\% of sensor readings according to this metric. A 6.14\% decrease to the mean percentage error in the predicted properties is also achieved compared to making predictions based on all available sensor readings.
\end{abstract}

\input{preface}

\tableofcontents                %% Mandatory
\listoftables                   %% Mandatory if thesis has tables
\listoffigures                  %% Mandatory if thesis has figures
%% Any other lists should come here, i.e. definitions, glossary, list of symbols etc.

\input{acknowledgements}

\chapter[Dedication]{} %% Optional
Dedicated to my wife Jill and daughter Jade.

\mainmatter % Now regular page numbering begins.

\input{1chapter1}
\input{2chapter2}
\input{3chapter3}
\input{4chapter4}
\input{5chapter5}
\input{6chapter6}
%\input{report}

\bibliographystyle{plain}
\bibliography{allbibs}

\appendix
%\input{software}
\input{hardware}

\backmatter

%% Indices come here if you have them.

\end{document}
\endinput
