\chapter{Preface} % Manditory if any of the conditions are met
The thesis is based on work conducted jointly between the Sensorimotor Systems and IMAGER laboratories with professors Dinesh K. Pai and Michiel van de Panne, and students Daniel Troniak and Chuan Zhu of the University of British Columbia. This section outlines the contributions of each of the above individuals. Students worked under guidance of the professors.  

\paragraph{Software}
Mr. Troniak wrote the C++ software framework for the control, perception and user interface of the robot, and developed MATLAB scripts in support of the design and usage of the TVR algorithm. Mr. Zhu implemented a MATLAB framework for data analysis, processing and figure generation, including the final implementation of the TVR algorithm. 

\paragraph{Experiments} 
Mr. Troniak designed and built the environment for the robot, collected all data with the robot, maintained and configured the hardware, and designed and executed the robot motion. Dr. Pai supported the purchase and maintenance of the robotic hardware. Mr. Zhu analyzed preliminary data collected during various manipulations to help Mr. Troniak determine successful motion trajectories. Data analysis, discussion, and final figure generation was accomplished collaboratively.

\paragraph{Algorithms}
Dr. van de Panne designed the TVR algorithm and proposed the usage of the PLS algorithm. Mr. Troniak validated the effectiveness of the TVR algorithm on data collected during manipulations. Mr. Zhu performed similar validation on the usage of the PLS algorithm, as well as designed the scheme for realtime predictions using PLS.

\paragraph{Writing}
Chapters \ref{chap3}, and \ref{chap5} of the thesis are adapted from a paper submitted to the 2014 IEEE International Conference on Robotics and Automation (D Troniak, C Zhu, D Pai, M van de Panne, Structured Predictions from Sensorimotor Data during Non-prehensile Manipulation, ICRA 2014, submission \#1515, 15 September 2013). Mr. Troniak performed and wrote the background literature review and drafted the remainder of the manuscript, which was then improved and prepared for submission by Dr. van de Panne.

\paragraph{Figures}
Figure \ref{fig:eg_sensor_trace} was generated by Mr. Troniak. Part of Figure \ref{fig:topple} was designed by Dr. van de Panne and is present in the paper submitted to ICRA 2014. 
%TODO
Figures from Chapter~\ref{chap5} were developed collaboratively and are also present in the paper. Figures borrowed from the literature include appropriate references within the caption.
